Atomify - a live LAMMPS visualizer

### 2017-05-10 ms.tex
Added figure, wrote some more

The work in this paper is focused on LAMMPS.
Although the software is well-documented and enables the user to perform
advanced simulations, it can be demanding to learn how to use it well.
A LAMMPS simulation is performed by a series of commands which are executed one by
one.
Physical quantities such as energy, temperature and stress can be measured by
adding \textit{computes}.
The time evolution of these quantities can in turn be saved to file for plotting
and further analysis.
Particle trajectories can be saved to a file and visualized by tools such as
VMD\cite{Humphrey1996Vmd} or OVITO\cite{Stukowski2009Visualization}.
The lack of an intuitive GUI and visualization\footnote{GUI is listed as a non-feature on the web page:
\url{http://lammps.sandia.gov/non_features.html}.} makes the workflow somewhat
complicated.

@@

TODO: Write about existing GUI software such as Materials Studio and MacroModel \\

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{flowchart.png}
	\caption{The typical workflow when using LAMMPS. a) A LAMMPS script is written using a text editor. b) Simulation is performed from the terminal. c) and d) visualization and analysis is performed using Ovito and \url{http://bit.ly/logplotter}. Atomify simplifies the workflow by combining all steps into one application. }
	\label{fig:flowchart}
\end{figure}

@@

Atomify provides a lightweight script editor with syntax highlighting and line
numbers.
% TODO add built-in help and autocompletion to Atomify and mention here
Multiple files can be open at the same time and persist across working sessions.
The script can easily be restarted by clicking the button in the GUI or using
the default shortcut (\keys{Ctrl+R} in Windows and Linux or \keys{\cmd+R} in macOS).
This enables you to quickly do changes in the script and see the changes immediately.

Certain Atomify-commands can be added to the comments of a regular LAMMPS script
to modify properties of the GUI. All these start with \keys{\texttt{\#}} so that
LAMMPS ignores them as they are interpreted as comments.
For instance, the color and size of the atoms can be set based on the type by
the command \keys{\texttt{\#/atom}}.
To set the color to white and radius to 2.0 for particle type 1,
the user can apply the command \keys{\texttt{\#/atom 1 2.0 \#ffffff}}. Certain built-in atom
types have default values and can be accessed with \keys{\texttt{\#/atom 1 oxygen}} for simplicity.
The initial camera position can also be set similarily.
For a full list of the available commands, see the Atomify documentation.

TODO: Add link to documentation

@@

Atomify provides live visualization of particles in a simulation in its 3D view.
The particles can be colored based on properties such as atom type,
velocity, and force.
The coloring scheme can be changed while the simulation is running by hovering
the mouse cursor over groups in the dashboard.
Light colors and strength can also be changed in the dashboard.
This will be described in further detail in the following section.

@@

In LAMMPS, computes are physical quantities calculated based on the running
simulation, such as energy, temperature, and stress.
% TODO better word for measures?
These are typically macroscopic measures derived from the microscopic
interactions between the atoms.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{final_billboard.png}
	\caption{Features}
	\label{fig:gui}
\end{figure}

>>>>

The work in this paper is focused on LAMMPS.
Although the software is well-documented and enables the user to perform
advanced simulations, it can be demanding to learn how to use it well.
A LAMMPS simulation is performed by a series of commands which are executed one by
one.
Physical quantities such as energy, temperature and stress can be measured by
adding \textit{computes}.
The time evolution of these quantities can in turn be saved to file for plotting
and further analysis.
Particle trajectories can be saved to a file and visualized by tools such as
VMD\cite{Humphrey1996Vmd} or OVITO\cite{Stukowski2009Visualization}.
The lack of an intuitive GUI and visualization\footnote{GUI is listed as a non-feature on the web page:
\url{http://lammps.sandia.gov/non_features.html}.} makes the workflow somewhat
complicated, see figure \ref{fig:flowchart}.

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{flowchart.png}
	\caption{The typical workflow when using LAMMPS consists of multiple steps using several applications. a) A LAMMPS script is written using a text editor. b) Simulation is performed from the terminal. c) and d) visualization and analysis is performed using Ovito and \url{http://bit.ly/logplotter}. Atomify simplifies the workflow by combining all steps into a single application. }
	\label{fig:flowchart}
\end{figure}

@@

@@

@@

Atomify provides a lightweight script editor with syntax highlighting and line
numbers.
% TODO add built-in help and autocompletion to Atomify and mention here
Multiple files can be open at the same time and the state persist across working sessions.
The script can easily be restarted by clicking the button in the GUI or using
the default shortcut (\keys{Ctrl+R} in Windows and Linux or \keys{\cmd+R} in macOS).
This enables you to quickly do changes in the script and see the changes immediately.

Certain Atomify-commands can be added to the comments of a regular LAMMPS script
to modify properties of the GUI. All these start with \keys{\texttt{\#}} so that
LAMMPS ignores them as they are interpreted as comments.
For instance, the color and size of the atoms can be set based on the type by
the command \keys{\texttt{\#/atom}}.
To set the color to white and radius to 2.0 for particle type 1,
the user can apply the command \keys{\texttt{\#/atom 1 2.0 \#ffffff}}. Certain built-in atom
types have default values and can be accessed with \keys{\texttt{\#/atom 1 oxygen}} for simplicity.
The initial camera position can also be set similarily.
For a full list of the available commands, see the Atomify documentation\cite{atomifydocumentation}.

@@

Atomify provides live visualization of particles in a simulation in its 3D view. 
The particles can be colored based on any produced per-atom quantity in LAMMPS such as
atom type, velocity, force or a variable formula.
The coloring scheme can be changed while the simulation is running by hovering
the mouse cursor over groups in the dashboard.
Light colors and strength can also be changed in the dashboard.
This will be described in further detail in the following section.

@@

In LAMMPS, computes are physical quantities calculated based on the running
simulation, such as energy, temperature and stress.
% TODO better word for measures?
These are typically macroscopic measures derived from the microscopic
interactions between the atoms.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.3\textwidth]{dashboard.png}
	\caption{Dashboard}
	\label{fig:gui}
\end{figure}
